\fontfamily{phv}\selectfont{Causal Discovery methods for Effective Connectivity in Human Brains}

### 2014-01-16 docs/paper/paper final.tex
Cleanup in 'Method'.

\paragraph{Stability}
The graph resulting from the PC algorithm depends on the order of the nodes\cite[p.120-122]{spirtes2000}.
It is therefore questionable whether strong conclusions about individual edges can be drawn from the output, as such an edge might not have appeared if the input was ordered differently.
As such, it seems useful to calculate a chance of a connection appearing.
When running the experiments, we have applied the algorithms multiple times using random permutations of the nodes to obtain stability selection as described by Meinshausen and B{\"u}hlmann \cite{meinshausen2010}.
These results have been avaraged to conclude the certainty of the found edges.

@@

% Implementation
We have implemented standard PC and EMS-PC in Matlab using the Bayesian Network Toolbox for Matlab.
% TODO: Finish this.

@@

%- Measuring whole-brain structural and directional consistency of PC - frequency of occurrences within subjects

We obtained structure and causal graphs for each subject with the previous procedure.
For each graph, we express the level of consistency of an edge in the frequency of occurrence of that type of edge over all permutations.
The frequency ratio per is then used to create the whole brain image for structure and directionality.
We compared the generated functional graph with the Bayesian inference approach on finding structural connectivity from Diffusion Weighted Imaging (DWI) proposed by Hinne et al.\cite{hinne2013} on the same data using edge consistency.

%Markov equivalence class size (nr. of latent variables / bidirectional arrows) within subjects and Markov class Asymmetry
Since standard PC and EMS-PC return Markov equivalence classes, we record the size of the solution space per subject.
Different permutations of nodes may yield different directionality for the same edge, including bi-directional relationships.
Of specific interest are consistent non-symmetrical orientations, as one-directional arrows imply a high-confidence causal relation.
Bi-directional edges imply an unseen common confounder.
We obtain uni-directional edges by subtracting the symmetrical bi-directional components from the causal graph.

@@

%- Comparison of structure (functional connectivity) of PC (both. they're identical in structure) with DWI results by Max et al. -> verplaatst naar het stukje aan het begin, over structure PC

\section{Results}\label{sec:results}

The results from the agorithms are presented in the form of matrices.
A connection from region $n$ to region $m$ is represented by point $(n,m)$ in the matrix.
Higher values represent strong certainties of the presence of a connection.
In such a matrix, indices 1-45 represent the brain regions of the left hemisphere, regions 45-90 represent the right hemisphere, and regions 91-116 represent regions in the cerebellum.

>>>>

@@

% Implementation
We have implemented standard PC and EMS-PC in Matlab using statistical tests and ideas from the Bayesian Network Toolbox for Matlab.\footnote{Bayesian Network Toolbox for Matlab: https://github.com/bayesnet/bnt}
This toolbox also includes an implementation for standard PC, however to gain better insight in the workings and to have greater flexibility in application, we have implemented standard PC separately.
The Matlab scripts used for the results are made freely available online.\footnote{Matlab implementation of EMS-PC: https://github.com/tomderuijter/EMS-PC}

@@

% Structural and Directional Consistency
% DWI Comparison
We obtained structure and causal graphs for each subject with the previous procedure.
For each graph, we express the level of consistency of an edge in the frequency of occurrence of that type of edge over all permutations.
The frequency ratio per is then used to create the whole brain image for structure and directionality.
We compared the generated functional graph with the Bayesian inference approach on finding structural connectivity from Diffusion Weighted Imaging (DWI) proposed by Hinne et al.\cite{hinne2013} on the same data using edge consistency.

%Markov equivalence class size and asymmetry
Since standard PC and EMS-PC return Markov equivalence classes, we record the size of the solution space per subject.
Different permutations of nodes may yield different directionality for the same edge, including bi-directional relationships.
Of specific interest are consistent non-symmetrical orientations, as one-directional arrows imply a high-confidence causal relation.
Bi-directional edges imply an unseen common confounder.
We obtain uni-directional edges by subtracting the symmetrical bi-directional components from the causal graph.

@@

\section{Results}\label{sec:results}
The results from the algorithms are presented in the form of matrices.
A connection from region $n$ to region $m$ is represented by point $(n,m)$ in the matrix.
Higher values represent strong certainties of the presence of a connection.
In such a matrix, indices 1-45 represent the brain regions of the left hemisphere, regions 45-90 represent the right hemisphere, and regions 91-116 represent regions in the cerebellum.
